---
title: "FA 14 Group 2"
output: html_document
date: "2022-11-01"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 01

### Reading the data

```{r}
library(GAD)

dat <- read.csv("https://raw.githubusercontent.com/tmatis12/datafiles/main/PowderProduction.csv")
```

### Item a

The equation for a three factor analysis is the following:

$$
y_{ijlk}=\mu+\alpha_i+\beta_j+\gamma_l+\alpha\beta_{ij}+\alpha\gamma_{il}+\beta\gamma_{jl}+\alpha\beta\gamma_{ijl}+\epsilon_{ijlk}
$$

### Item b

#### Hypothesis

The hypothesis that we are trying to reject are the following (not in order of relevance):

$$
Ho: \alpha\beta\gamma_{ijl}= 0\\
Ha: \alpha\beta\gamma_{ijl}\neq 0
$$

$$
Ho: \alpha\beta_{ij}= 0\\
Ha: \alpha\beta_{ij}\neq 0
$$

$$
Ho: \beta\gamma_{jl}= 0\\
Ha: \beta\gamma_{jl}\neq 0
$$

$$
Ho: \alpha\gamma_{il}= 0\\
Ha: \alpha\gamma_{il}\neq 0
$$

$$
Ho: \alpha_{i}= 0\\
Ha: \alpha_{i}\neq 0
$$

$$
Ho: \beta_{j}= 0\\
Ha: \beta_{j}\neq 0
$$

$$
Ho: \gamma_{l}= 0 \\
Ha: \gamma_{l}\neq 0
$$

#### Preparing the data

```{r}
dat$Ammonium <- as.fixed(dat$Ammonium)
dat$StirRate <- as.fixed(dat$StirRate)
dat$Temperature <- as.fixed(dat$Temperature)
```

#### Statistical Analysis

To test our hypothesis, let's perform the following:

##### Testing the three factor interaction

$$
Ho: \alpha\beta\gamma_{ijl}= 0\\
Ha: \alpha\beta\gamma_{ijl}\neq 0
$$

```{r}
model.aov <- aov(Density~Ammonium+StirRate+Temperature+Ammonium*StirRate+Ammonium*Temperature+StirRate*Temperature+Ammonium*StirRate*Temperature, data=dat)
summary(model.aov)
```

From the results above, it we cannot reject Ho, since the p-value is greater than $\alpha$. Therefore, the interaction between the three factors is not significant.

Now, let's test the least significant effect in order to validate the two factors interaction by removing the three factor interaction.

##### Testing alpha and gamma interaction

$$
Ho: \alpha\gamma_{il}= 0\\
Ha: \alpha\gamma_{il}\neq 0
$$

```{r}
model.aov <- aov(Density~Ammonium+StirRate+Temperature+Ammonium*StirRate+Ammonium*Temperature+StirRate*Temperature, data=dat)
summary(model.aov)
```

Since the p-value is greater than $\alpha$, we cannot reject Ho. Therefore, we need to check the next least significant interaction by removing alpha gamma interation.

##### Testing beta gamma interaction

$$
Ho: \beta\gamma_{jl}= 0\\
Ha: \beta\gamma_{jl}\neq 0
$$

```{r}
model.aov <- aov(Density~Ammonium+StirRate+Temperature+Ammonium*StirRate+StirRate*Temperature, data=dat)
summary(model.aov)
```

Since the p-value is greater than $\alpha$, we cannot reject Ho. Therefore, we need to test the last two factor interaction to see its relevance by removing beta gamma interaction.

##### Testing alpha beta interaction 

$$
Ho: \alpha\beta_{ij}= 0\\
Ha: \alpha\beta_{ij}\neq 0
$$

```{r}
model.aov <- aov(Density~Ammonium+StirRate+Temperature+Ammonium*StirRate, data=dat)
summary(model.aov)
```

Since the p-value of the last two factor interaction is lesser than $\alpha$ ($p-value=0.022<\alpha=0.05$), we can reject Ho and consider that there is a significant interaction between the Ammonium percentage and the Stir Rate.

Since there is an interaction going on, we can check the plot by

```{r}
interaction.plot(dat$StirRate,dat$Ammonium,dat$Density)

```

From the plot, since the curves are not parallel, we can confirm the interaction between the two factors.

## Question 02

### Part a

$$
Ho: \alpha\beta_{ij}=0 \\
Ha: \alpha\beta_{ij}\neq0
$$

```{r}
library(GAD)
position <- c(rep(1,9),rep(2,9))
temperature <- c(rep(seq(800,850,25),6))
response <- c(570,	1063,	565,
              565,	1080,	510,
              583,	1043,	590,
              528,	988,	526,
              547,	1026,	538,
              521,	1004,	532)
position <- as.fixed(position)
temperature <- as.fixed(temperature)
dat <- data.frame(position,temperature,response)
dat
model1 <- aov(response~position+temperature+position*temperature,data=dat)
gad(model1)
interaction.plot(temperature,position,response)
```

P value of the interaction is 0.427110 which is larger than $\alpha = 0.05$ level of significance, so we don't reject H0 hence there is no two factors interaction present. Also p value for the position and temperature is 0.001762 and 3.25e-14 respectively.

### Part b

$$
Ho: \alpha\beta_{ij}=0 \\
Ha: \alpha\beta_{ij}\neq0
$$

```{r}
library(GAD)
position <- c(rep(1,9),rep(2,9))
temperature <- c(rep(seq(800,850,25),6))
response <- c(570,	1063,	565,
              565,	1080,	510,
              583,	1043,	590,
              528,	988,	526,
              547,	1026,	538,
              521,	1004,	532)
position <- as.random(position)
temperature <- as.random(temperature)
dat2 <- data.frame(position,temperature,response)
dat2
model2 <- aov(response~position+temperature+position*temperature,data=dat2)
gad(model2)
```

P value of the interaction is 0.4271101 which is larger than $\alpha = 0.05$ level of significance, so we don't reject H0 hence there is no two factors interaction present. Also p value for the position and temperature is 0.0526583 and 0.0008647 respectively.

### Part c

$$
Ho: \alpha\beta_{ij}=0 \\
Ha: \alpha\beta_{ij}\neq0
$$

```{r}
library(GAD)
position <- c(rep(1,9),rep(2,9))
temperature <- c(rep(seq(800,850,25),6))
response <- c(570,	1063,	565,
              565,	1080,	510,
              583,	1043,	590,
              528,	988,	526,
              547,	1026,	538,
              521,	1004,	532)
position <- as.fixed(position)
temperature <- as.random(temperature)
dat3 <- data.frame(position,temperature,response)
dat3
model3 <- aov(response~position+temperature+position*temperature,data=dat3)
gad(model3)
```

P value of the interaction is 0.427110 which is larger than $\alpha = 0.05$ level of significance, so we don't reject H0 hence there is no two factors interaction present. Also p value for the position and temperature is 0.05266 and 3.25e-14 respectively.

### Part d

Comments on p values: 

The interaction term is not significant for all the part A, B and C.

When both factors are fixed, position and temperature both are significant at alpha= 0.01 and 0.001 level of significance.

When both factors are random, position and temperature both are significant at alpha= 0.1 and 0.001 level of significance.

When the factors have mixed effect, position and temperature both are significant at alpha= 0.1 and 0.001 level of significance.

But, it is possible to see a difference in the p-values of the main effects:

For fixed models, we are getting different p-values when compared with the random models. However, for mixed models, we are getting smaller p-values for the main effects. 

Temperature p-value is the same for Fixed and Mixed Models, but different in the random

Position p-value is the same for mixed and Random but different for fixed
 
