---
title: "Flip Assignment 12"
author: "Yashwanth Dommaraju, Gustavo Marin Paulon, Md Ariful Haque Miah"
date: "2022-10-13"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data

## Question 1 


For the Latin Square become valid, the Latin letters should not repeat in each row and column.
So, since it is a two-source of nuisance variability that is exactly the case, the Latin square is valid.

## Question 2

Since it is a two-block Latin Square (Latin Square), let's consider the following equation: 

$$
y_{ijk}=\mu+\tau_i+\beta_j+\alpha_k+\epsilon_{ijk}
$$

Where $\tau_i$ is the treatment effect, $\beta_j$ is the nuisance variability of the Batch (block one), $\alpha_k$ is the nuisance variability of the Days (block two) and $\epsilon_{ijk}$ is the random error.

## Question 3

The hypothesis of the statistics test is:

$$
H_o: \tau_i = 0 \;for \;all \;i \\
H_a: \tau_i \neq 0 \;for \;some \;i \;(i=1,2,3,4,5)
$$

```{r}
library(GAD)

batch <- c(rep(1,5),rep(2,5),rep(3,5),rep(4,5),rep(5,5))
day <- c(rep(seq(1,5),5))

ing <- c("A","B","D","C","E",
         "C","E","A","D","B",
         "B","A","C","E","D",
         "D","C","E","B","A",
         "E","D","B","A","C")

obs <- c(8,7,1,7,3,
         11,2,7,3,8,
         4,9,10,1,5,
         6,8,6,6,10,
         4,2,3,8,8)

batch <- as.fixed(batch)
day <- as.fixed(day)

ing <- as.fixed(ing)

data <- data.frame(batch,day,ing,obs)

model <- aov(obs~batch+day+ing, data=data)

summary(model)
```

Conclusion: Since the p-value of ingredients is lesser than alpha ($p-value_{ing}=0.000488 < \alpha$) we can reject $H_o$.

Therefore, changing the ingredient can significantly change the reaction time of chemical process, because the treatment effect is different than zero.



