---
title: "HW-Week 7"
author: "Md Ariful Haque Miah"
date: "10/13/2022"
output:
  html_document:
    df_print: paged
---
Problem 4.3

```{r}
library(GAD)
chemical<-c(rep(1,5),rep(2,5),rep(3,5),rep(4,5))
bolt<-c(rep(seq(1,5),4))
obs <- c(73,68,74,71,67,73,67,75,72,70,75,68,78,73,68,73,71,75,75,69)
chemical<-as.fixed(chemical)
bolt<-as.fixed(bolt)
obs <- as.numeric(obs)
dat <- data.frame(chemical,bolt,obs)
dat
str(dat)
model<-lm(obs~chemical+bolt,data=dat) 
gad(model)
```

Hypothesis:

$H_0:\tau_i = 0$ for all i

$H_a:\tau_i \neq 0$ for some i

Model Equation:

$y_{ij}=\mu+\tau_i+\beta_j+e_{ij}$

where $\mu$ = grand mean, $\tau_i$ = Chemical effect,
$\beta_j$ = Bolt (block), $e_{ij}$ = random error ~ N(0,$\sigma^2$)

P value (0.1211) > $\alpha$ (0.05) hence we fail to reject $H_0$.

There is no difference among the four chemical agents at $\alpha$ = 0.05 level.

Problem 4.16

Calculate model parameters $\tau_i$:
Formula: $\tau_i$ = $\mu_i-\mu$

```{r}
obs <- c(73,68,74,71,67,73,67,75,72,70,75,68,78,73,68,73,71,75,75,69)
mean(obs)
tau_1 <- mean(obs[1:5])-mean(obs)
tau_1
tau_2 <- mean(obs[6:10])-mean(obs)
tau_2
tau_3 <- mean(obs[11:15])-mean(obs)
tau_3
tau_4 <- mean(obs[16:20])-mean(obs)
tau_4
```

$\tau_1$ = -1.15, $\tau_2$ = -0.35, $\tau_3$ = 0.65, $\tau_4$ = 0.85

Calculate model parameters $\beta_j$:
Formula: $\beta_j$ = $\mu_j-\mu$

```{r}
obs <- c(73,68,74,71,67,73,67,75,72,70,75,68,78,73,68,73,71,75,75,69)
beta_1 <- mean(c(73,73,75,73))-mean(obs)
beta_1
beta_2 <- mean(c(68,67,68,71))-mean(obs)
beta_2
beta_3 <- mean(c(74,75,78,75))-mean(obs)
beta_3
beta_4 <- mean(c(71,72,73,75))-mean(obs)
beta_4
beta_5 <- mean(c(67,70,68,69))-mean(obs)
beta_5
```

$\beta_1$ = 1.75, $\beta_2$ = -3.25, $\beta_3$ = 3.75, $\beta_4$ = 1, $\beta_5$ = -3.25

Problem 4.22

```{r}
library(GAD)
batch <- c(rep(1,5),rep(2,5),rep(3,5),rep(4,5),rep(5,5))
day <- c(rep(seq(1,5),5))
ingr <- c("A","B","D","C","E",
         "C","E","A","D","B",
         "B","A","C","E","D",
         "D","C","E","B","A",
         "E","D","B","A","C")
obs <- c(8,7,1,7,3,
         11,2,7,3,8,
         4,9,10,1,5,
         6,8,6,6,10,
         4,2,3,8,8)
batch <- as.fixed(batch)
day <- as.fixed(day)
ingr <- as.fixed(ingr)
obs <- as.numeric(obs)
data <- data.frame(batch,day,ingr,obs)
data
str(data)
model <- aov(obs~batch+day+ingr, data=data)
summary(model)
```

Hypothesis 
$H_0$: $\tau_i$ = 0 for all i

$H_a$: $\tau_i$ $\neq$ 0 for some i, i=1,2,3,4,5

Model equation: $X_{ijk}$ = $\mu + \tau_i + \beta_j + \alpha_k + e_{ijk}$

where, $\mu$=grand mean, $\tau_i$=Ingredients effect, $\beta_j$ = Batch (block 1)
$\alpha_k$ = Day (block 2), $e_{ijk}$ = random error ~N(0,$\sigma^2$)

Conclusions:

Hence the p value (0.000488) of ingredients is highly significant and less than
$\alpha$=0.05 so we reject the null hypothesis $H_0$.

Therefore the effect of five different ingredients (A,B,C,D,E) on the reaction time
of a chemical process is significantly different. 

Source Code

```{r eval=FALSE}
library(GAD)
chemical<-c(rep(1,5),rep(2,5),rep(3,5),rep(4,5))
bolt<-c(rep(seq(1,5),4))
obs <- c(73,68,74,71,67,73,67,75,72,70,75,68,78,73,68,73,71,75,75,69)
chemical<-as.fixed(chemical)
bolt<-as.fixed(bolt)
obs <- as.numeric(obs)
dat <- data.frame(chemical,bolt,obs)
dat
str(dat)
model<-lm(obs~chemical+bolt,data=dat) 
gad(model)

obs <- c(73,68,74,71,67,73,67,75,72,70,75,68,78,73,68,73,71,75,75,69)
mean(obs)
tau_1 <- mean(obs[1:5])-mean(obs)
tau_1
tau_2 <- mean(obs[6:10])-mean(obs)
tau_2
tau_3 <- mean(obs[11:15])-mean(obs)
tau_3
tau_4 <- mean(obs[16:20])-mean(obs)
tau_4

obs <- c(73,68,74,71,67,73,67,75,72,70,75,68,78,73,68,73,71,75,75,69)
beta_1 <- mean(c(73,73,75,73))-mean(obs)
beta_1
beta_2 <- mean(c(68,67,68,71))-mean(obs)
beta_2
beta_3 <- mean(c(74,75,78,75))-mean(obs)
beta_3
beta_4 <- mean(c(71,72,73,75))-mean(obs)
beta_4
beta_5 <- mean(c(67,70,68,69))-mean(obs)
beta_5

library(GAD)
batch <- c(rep(1,5),rep(2,5),rep(3,5),rep(4,5),rep(5,5))
day <- c(rep(seq(1,5),5))
ingr <- c("A","B","D","C","E",
         "C","E","A","D","B",
         "B","A","C","E","D",
         "D","C","E","B","A",
         "E","D","B","A","C")
obs <- c(8,7,1,7,3,
         11,2,7,3,8,
         4,9,10,1,5,
         6,8,6,6,10,
         4,2,3,8,8)
batch <- as.fixed(batch)
day <- as.fixed(day)
ingr <- as.fixed(ingr)
obs <- as.numeric(obs)
data <- data.frame(batch,day,ingr,obs)
data
str(data)
model <- aov(obs~batch+day+ingr, data=data)
summary(model)
```

