---
title: "HW Week 14"
author: "Md Ariful Haque Miah, Ayoola Ayodegi"
date: "2022-12-04"
output: html_document
---
Problem 14.3

```{r}
library(GAD)
machine <- c(rep(1,8),rep(2,8),rep(3,8))
spindle <- rep(c(rep(1,4),rep(2,4)),3)
observation <- c(12,9,11,12,8,9,10,8,14,15,13,14,12,10,11,13,14,10,12,11,16,15,15,14)
machine <- as.fixed(machine)
spindle <- as.random(spindle)
dat <- data.frame(machine,spindle,observation)
dat
model <- lm(observation~machine+spindle%in%machine,data=dat)
gad(model)
```

Model Equation:
Y-ijk = mu + alpha_i + beta_j(i) + e_ijk, where spindle is nested within machine.
Hypothesis Test:

Null H_0: alpha_i = 0 for all i
Alternative H_a: alpha_i != 0 for some i

Null H_0: beta_j(i) = 0 for all i,j
Alternative H_a: beta_j(i) != 0 for some i,j

From the summary we see that, the factor machine is not significant since P value (0.2915630) is higher than
$\alpha = 0.05$ level of significance thus we fail to reject $H_0$.
On the other hand, the factor spindle is nested within the factor machine is significant since P value (0.0004428) is lower than $\alpha = 0.05$ level of significance thus we reject $H_0$.

Source Code:
```{r eval=FALSE}
library(GAD)
machine <- c(rep(1,8),rep(2,8),rep(3,8))
spindle <- rep(c(rep(1,4),rep(2,4)),3)
observation <- c(12,9,11,12,8,9,10,8,14,15,13,14,12,10,11,13,14,10,12,11,16,15,15,14)
machine <- as.fixed(machine)
spindle <- as.random(spindle)
dat <- data.frame(machine,spindle,observation)
dat
model <- lm(observation~machine+spindle%in%machine)
gad(model)
```


