---
title: "FA16"
output: html_document
date: "2022-11-08"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Input Data

```{r}
library(DoE.base)
A <- c(-1,1)
B <- c(rep(-1,2),rep(1,2))
C <- c(rep(-1,4),rep(1,4))
D <- c(rep(-1,8),rep(1,8))
obs <- c(12,18,13,16,17,15,20,15,10,25,13,24,19,21,17,23)
dat <- data.frame(A,B,C,D,obs)
dat
mod <- lm(obs~A*B*C*D,data=dat)
halfnormal(mod)
```

## Item a

From the plots, the factors A, D, A:C, and A:D are significant since they do not follow a normal line. In other words, they do not behave like a random error.

## Item b

### Hypothesis

$$
H_o: \alpha\delta_{im}=0 \\ H_a: \alpha\delta_{im}\neq0 
$$

$$
H_o: \alpha\gamma_{il}=0 \\ H_a: \alpha\gamma_{il}\neq0 
$$

Where, $\alpha$ is the main effect of A, $\gamma$ is the main effect of C, $\delta$ is the main effect of D.

### Hypothesis Testing

```{r}
mod2 <- lm(obs~A+C+D+A*C+A*D,data=dat)

summary(mod2)
```

From the test, we can assure now that the terms that remain the linear model equation, whiche are only five out of sixteen (A,C,D,A:C,A:D), are significant. All the p-values are lesser than alpha ($\alpha=0.05$).

```{r}
mod3 <- aov(obs~A+C+D+A*C+A*D,data=dat)

summary(mod3)
```

