---
title: "HW Week 11"
author: "Md Ariful Haque Miah, Ayodegi"
date: "11/13/2022"
output:
  html_document:
    df_print: paged
editor_options: 
  markdown: 
    wrap: 72
---

Problem 6.8

```{r}
library(DoE.base)
time<-c(rep(-1, 12),rep(1, 12) )
cul_med<-c(rep(-1, 2), rep(1, 2))
obs <- c(21,22,25,26,23,28,24,25,20,26,29,27,37,39,31,34,38,38,29,33,35,36,30,35)
time <- as.factor(time)
cul_med <- as.factor(cul_med)
dat <- data.frame(time,cul_med,obs)
dat
model <- aov(obs~time*cul_med,data=dat)
summary(model) 
plot(model)
```

Hypothesis Test:

$H_0: \alpha_i = 0$

$H_a: \alpha_i \neq 0$

$H_0: \beta_j\ = 0$

$H_a: \beta_j \neq 0$

$H_0: \alpha\beta_{ij}=0$

$H_a: \alpha\beta_{ij}\neq0$

Only the factors Time and the interaction effects (Time:Culture Medium)
are significant as the p value is less than $\alpha=0.001$ level of
significance so for these factors we reject $H_0$ and the factor Culture
medium is not significant as the p value is larger than $\alpha=0.05$
level of significance so for this factors we fail to reject $H_0$. Also
Residual vs Fitted plot shows that the variance is constant and the
errors are normally distributed and thus the model is adequate.

Problem 6.12 (a) Estimate the factor effects

```{r}
A <- c(rep(-1,4),rep(1,4))
B <- c(rep(-1,8),rep(1,8))
obs <- c(14.037,16.165,13.972,13.907,13.880,13.860,14.032,13.914,14.821,14.757,14.843,14.878,14.888,14.921,14.415,14.932)
dat <- data.frame(A,B,obs)
dat
mod <- lm(obs~A*B,data=dat)
coef(mod)




A <- as.factor(A)
B <- as.factor(B)
dat <- data.frame(A,B,obs)
dat
```
**Multiply the value of each factor's coefficients by 2, we will get the corresponding factor effects.**

Finding Corner Points:

```{r}
one <- sum(dat$obs[1:4])
one
a <- sum(dat$obs[5:8])
a
b <- sum(dat$obs[9:12])
b
ab <- sum(dat$obs[13:16])
ab
n <- c(4)
fact_A <- 2*(a+ab-b-one)/(4*n)
fact_A
fact_B <- 2*(b+ab-a-one)/(4*n)
fact_B
fact_AB <- 2*(ab+one-a-b)/(4*n)
fact_AB
```

Therefore, The factor Effect A = -0.31725, B = 0.586, Interaction effect
AB = 0.2815

(b) Analysis of variance

```{r}
A <- c(rep(-1,4),rep(1,4))
B <- c(rep(-1,8),rep(1,8))
obs <- c(14.037,16.165,13.972,13.907,13.880,13.860,14.032,13.914,14.821,14.757,14.843,14.878,14.888,14.921,14.415,14.932)
A <- as.factor(A)
B <- as.factor(B)
dat <- data.frame(A,B,obs)
dat
model <- aov(obs~A*B,data=dat)
summary(model)
```

Hypothesis Test: $H_0: \alpha_i = 0$

$H_a: \alpha_i \neq 0$

$H_0: \beta_j\ = 0$

$H_a: \beta_j \neq 0$

$H_0: \alpha\beta_{ij}=0$

$H_a: \alpha\beta_{ij}\neq0$

p values of the main effect A and the interaction effect (A:B) is
greater than $\alpha = 0.05$ level of significance. Hence the main
effect A and the interaction effect (A:B) are not significant as we fail
to reject $H_0$. On the other hand p value of the main effect B is less
than $\alpha = 0.1$ level of significance. Hence the main effect B is
significant as we reject $H_0$.

(c) Regression Equation:

```{r}
A <- c(rep(-1,4),rep(1,4))
B <- c(rep(-1,8),rep(1,8))
obs <- c(14.037,16.165,13.972,13.907,13.880,13.860,14.032,13.914,14.821,14.757,14.843,14.878,14.888,14.921,14.415,14.932)
A <- as.factor(A)
B <- as.factor(B)
dat <- data.frame(A,B,obs)
dat
model <- lm(obs~A+B+A*B,data=dat)
summary(model)
```

y = $\beta_0 + \beta_2 *x_2 + \epsilon$

Hence, y = $14.5203 + 0.3045 * x_2 + \epsilon$

where $\beta_0$ = the intercept = the grand mean of all 16 observations
and the regression coefficient $\beta_2$ is one-half the corresponding
factor effect B estimate.

(d) The Residuals Analysis

```{r}
A <- c(rep(-1,4),rep(1,4))
B <- c(rep(-1,8),rep(1,8))
obs <- c(14.037,16.165,13.972,13.907,13.880,13.860,14.032,13.914,14.821,14.757,14.843,14.878,14.888,14.921,14.415,14.932)
A <- as.factor(A)
B <- as.factor(B)
dat <- data.frame(A,B,obs)
dat
model <- aov(obs~A*B,data=dat)
summary(model)
plot(model)
```

There is a significant concern in the Residuals vs Fitted Plot, which
shows that the variance is not constant. Hence the model is not
adequate. Though the normality assumptions are fairly satisfied in the
Normal Q-Q plot.

(e) The potential Outlier: Only the observation no. 2 might be the
    potential outlier according to the $\sqrt$(standardized residual)
    plot in part (d). Since There is 15 more observations left so only
    one outlier won't cause any effect to grow an epitaxial layer on
    polished silicon wafers. Also,the interaction effects are not significant, we can choose any main effect  to maximize the yield of the response vairable (epitaxal layer thickness).

Problem 6.21 (a)

```{r}
A <- c(rep(-1,7),rep(1,7))
B <- c(rep(-1,14),rep(1,14))
C <- c(rep(-1,28),rep(1,28))
D <- c(rep(-1,56),rep(1,56))
obs <- c(10,18,14,12.5,19,16,18.5,0,16.5,4.5,17.5,20.5,17.5,33,4,6,1,14.5,12,14,5,0,10,34,11,25.5,21.5,0,0,0,18.5,19.5,16,15,11,5,20.5,18,20,29.5,19,10,6.5,18.5,7.5,6,0,10,0,16.5,4.5,0,23.5,8,8,8,4.5,18,14.5,10,0,17.5,6,19.5,18,16,5.5,10,7,36,15,16,8.5,0,0.5,9,3,41.5,39,6.5,3.5,7,8.5,36,8,4.5,6.5,10,13,41,14,21.5,10.5,6.5,0,15.5,24,16,0,0,0,4.5,1,4,6.5,18,5,7,10,32.5,18.5,8)
A <- as.factor(A)
B <- as.factor(B)
C <- as.factor(C)
D <- as.factor(D)
dat <- data.frame(A,B,C,D,obs)
dat
model <- aov(obs~A*B*C*D,data=dat)
summary(model)
```

Hypothesis Test:

$H_0: \alpha_i = 0$

$H_a: \alpha_i \neq 0$

$H_0: \beta_j\ = 0$

$H_a: \beta_j \neq 0$

$H_0: \gamma_k\ = 0$

$H_a: \gamma_k \neq 0$

$H_0: \delta_l\ = 0$

$H_a: \delta_l \neq 0$

$H_0: \alpha\beta_{ij}=0$

$H_a: \alpha\beta_{ij}\neq0$

$H_0: \alpha\gamma_{ik}=0$

$H_a: \alpha\gamma_{ik}\neq0$

$H_0: \alpha\delta_{il}=0$

$H_a: \alpha\delta_{il}\neq0$

.....

.....

.....

$H_0: \alpha\beta\gamma\delta_{ijkl}=0$

$H_a: \alpha\beta\gamma\delta_{ijkl}\neq0$

From the anova model summary and analyzing the p value we see that None
of the interaction terms are significant but the main effect factors
"length of putt" and "the type of putter" are significant and these
factors significantly affect putting performance.

Problem 6.21 (b)

```{r}
plot(model)
```

Residual vs fitted plot shows that the variance is not constant and
hence the model is not adequate though the errors are assumed to be
normally distributed from the normal Q-Q plot.
Check boxcox

```{r}
library(MASS)
A <- c(rep(10,7),rep(30,7),rep(10,7),rep(30,7),rep(10,7),rep(30,7),rep(10,7),rep(30,7),rep(10,7),rep(30,7),rep(10,7),rep(30,7),rep(10,7),rep(30,7),rep(10,7),rep(30,7))
B <- c(rep(1,14),rep(2,14),rep(1,14),rep(2,14),rep(1,14),rep(2,14),rep(1,14),rep(2,14))
C <- c(rep(1,28),rep(2,28),rep(1,28),rep(2,28))
D <- c(rep(1,56),rep(2,56))
obs <- c(10,18,14,12.5,19,16,18.5,0,16.5,4.5,17.5,20.5,17.5,33,4,6,1,14.5,12,14,5,0,10,34,11,25.5,21.5,0,0,0,18.5,19.5,16,15,11,5,20.5,18,20,29.5,19,10,6.5,18.5,7.5,6,0,10,0,16.5,4.5,0,23.5,8,8,8,4.5,18,14.5,10,0,17.5,6,19.5,18,16,5.5,10,7,36,15,16,8.5,0,0.5,9,3,41.5,39,6.5,3.5,7,8.5,36,8,4.5,6.5,10,13,41,14,21.5,10.5,6.5,0,15.5,24,16,0,0,0,4.5,1,4,6.5,18,5,7,10,32.5,18.5,8)
model <- aov(obs~A*B*C*D)
summary(model)
plot(model)
boxcox(model)
```



## QUESTION 6.36

Getting in data

```{r,warning=FALSE,message=FALSE}
library(DoE.base)
A<- c(-1,1)
B<- c(rep(-1,2),rep(1,2))
C<- c(rep(-1,4),rep(1,4))
D<- c(rep(-1,8),rep(1,8))
obs<- c(1.92,11.28,1.09,5.75,2.13,9.53,1.03,5.35,1.6,11.73,1.16,4.68,2.16,9.11,1.07,5.3)
dat<- data.frame(A,B,C,D,obs)
dat
```

## 6.36) A

```{r}
model<-lm(obs~A*B*C*D,data=dat)
coef(model)
```

```{r}
halfnormal(model)
```

From the Half-normal plot,we can see that Factors A,B, A:B , and A:B:C are significant at alpha=0.5

```{r}
model2<-lm(obs~A+B+C+A*B+A*B*C,data=dat)
summary(model2)
```

To select our tentative model, we would consider the significant factors from our half-normal plots,which are factors A,B,A:B,A:B:C.

So therefore our tentative model can be written as

$Y_{i,j,k,l}=4.68062+3.16062\alpha _{i}-1.50187\beta_{j}-1.06938\alpha\beta_{ij}+0.34437\alpha\beta\gamma+\epsilon_{ijkl}$

## 6.36) B

```{r}
model2<- aov(obs~A+B+C+A*B+A*B*C,data =dat)
summary(model2)
```

```{r}
plot(model2)
```

From the Normal Q-Q plot, we can't assume Normality in the data since the data points in the plot doesn't appears to fall on a straight line

Also, from the Residuals vs fitted plot, we can't assume constant variance since all the residuals have varying spread, depleting that the variances are not equal.

## 6.36) C

Using a Log transformation on our data below

```{r}
logobs <- log(obs)
dat2 <- data.frame(A,B,C,D,logobs)
model3<- lm(logobs~A*B*C*D,data = dat2)
coef(model3)
```

```{r}
halfnormal(model3)
```

```{r}
model3<-lm(logobs~A+B+C+A:B:C,data=dat2)
summary(model3)
```

```{r}
model4<-aov(logobs~A+B+C+A:B:C,data=dat2)
summary(model4)

```

```{r,warning=FALSE,message=FALSE}
plot(model4)
```

After the Log transformation, we can observe that from the half-normal plots that only factors A, B, and A:B:C are significant and interactions of A:B seems insignificant as compared to pre-log transformation.

The confirmation of the half normal plot was double checked using ANOVA analysis that showed factors A,B, and A:B:C to be significant as shown above.

Based on the residuals plots, after the log transformation was performed we observed that

From the Normal Q-Q plot, we can not assume Normality in the data since the data points in the plot does-not appear to fall on a straight line

Also, from the Residuals vs fitted plot, we cannot assume constant variance since all the residuals have varying spread, depleting that the variances are not equal.

## 6.36) D

Fitting a model of the coded variables we have

$Y_{ijkl}=1.185417+0.812870\alpha _{i}-0.314278\beta_{j}+0.063434\alpha\beta\gamma+\epsilon_{{ijkl}}$



## QUESTION 6.39

Getting in the data we have

```{r}
library(DoE.base)
A<-c(-1,1,-1,1,-1,1,-1,1,-1,1,-1,1,-1,1,-1,1,-1,1,-1,1,-1,1,-1,1,-1,1,-1,1,-1,1,-1,1)
B<-c(-1,-1,1,1,-1,-1,1,1,-1,-1,1,1,-1,-1,1,1,-1,-1,1,1,-1,-1,1,1,-1,-1,1,1,-1,-1,1,1)
C<-c(-1,-1,-1,-1,1,1,1,1,-1,-1,-1,-1,1,1,1,1,-1,-1,-1,-1,1,1,1,1,-1,-1,-1,-1,1,1,1,1)
D<-c(-1,-1,-1,-1,-1,-1,-1,-1,1,1,1,1,1,1,1,1,-1,-1,-1,-1,-1,-1,-1,-1,1,1,1,1,1,1,1,1)
E<- c(-1,-1,-1,-1,-1,-1,-1,-1,-1,-1,-1,-1,-1,-1,-1,-1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1)
obs<- c(8.11,5.56,5.77,5.82,9.17,7.8,3.23,5.69,8.82,14.23,9.2,8.94,8.68,11.49,6.25,9.12,7.93,5,7.47,12,9.86,3.65,6.4,11.61,12.43,17.55,8.87,25.38,13.06,18.85,11.78,26.05)
dat<- data.frame(A,B,C,D,E,obs)
```

## 6.39) A

```{r}
model<-lm(obs~A*B*C*D*E,data=dat)
coef(model)
```

```{r}
halfnormal(model)
```

```{r}
summary(lm(obs~B+D+ E+ A:D+ A+ D:E+ B:E+ A:B+ A:B:E+ A:E+ A:D:E,data=dat))
```

```{r}
summary(model)
```

```{r}
model2<- aov(obs~A+B+D+E+A*B+A*D+A*E+B*E+D*E+A*B*E+A*D*E,data=dat)
summary(model2)
```

We can observe that from the half-normal plots that A,D,E,A:D,D:E,B:E,A:B,A:E,A:B:E,A:D:E are significant

The confirmation of the half normal plot was double checked using ANOVA analysis that showed that the above factors mentioned were significant

## 6.39) B

```{r}
plot(model2)
```

From the Normal Q-Q plot, we can't assume Normality in the data since the data points in the plot doesn't appears to fall on a straight line

Also, from the Residuals vs fitted plot, we can't assume constant variance since all the residuals have varying spread, depleting that the variances are not equal.

## 6.39) C

```{r}
A<- c(-1,1,-1,1,-1,1,-1,1,-1,1,-1,1,-1,1,-1,1,-1,1,-1,1,-1,1,-1,1,-1,1,-1,1,-1,1,-1,1)
B<- c(-1,-1,1,1,-1,-1,1,1,-1,-1,1,1,-1,-1,1,1,-1,-1,1,1,-1,-1,1,1,-1,-1,1,1,-1,-1,1,1)
D<- c(-1,-1,-1,-1,-1,-1,-1,-1,1,1,1,1,1,1,1,1,-1,-1,-1,-1,-1,-1,-1,-1,1,1,1,1,1,1,1,1)
E<- c(-1,-1,-1,-1,-1,-1,-1,-1,-1,-1,-1,-1,-1,-1,-1,-1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1)
obs<- c(8.11,5.56,5.77,5.82,9.17,7.8,3.23,5.69,8.82,14.23,9.2,8.94,8.68,11.49,6.25,9.12,7.93,5,7.47,12,9.86,3.65,6.4,11.61,12.43,17.55,8.87,25.38,13.06,18.85,11.78,26.05)
dat<- data.frame(A,B,D,E,obs)
```

```{r}
model4<- lm(obs~A*B*D*E,data =dat)
coef(model4)
```

```{r}
halfnormal(model4)
```

```{r}
summary(model4)
```

```{r}
model5<- aov(obs~A+B+D+E+A*B+A*D+A*E+B*E+D*E+A*B*E+A*D*E,data=dat)
summary(model5)
```

```{r}
plot(model5)
```

Factor C was dropped totally from our model initially because it seemed insignificant. After dropping factor C we realized that from the half-normal plots that A,D,E,A:D,D:E,B:E,A:B,A:E,A:B:E,A:D:E are significant

The confirmation of the half normal plot was double checked using ANOVA analysis that showed that the above factors mentioned were significant

Which was similar to the results gotten when factor C was included.

## 6.39) D

Fitting our model in a linear equation we can see that

$Y_{ijkl}=10.18031+1.61594\alpha _{i}+0.04344\beta_{j}+2.98844\gamma_{k}+1.66656\alpha\gamma+1.38969\gamma\delta+1.28344\beta\delta+1.23656\alpha\beta+1.02719\alpha\delta+1.18531\alpha\beta\delta+0.90156\alpha\gamma\delta$

To maximize the predicted response, the above linear model should be follows

Note

Because all the factors are of positive coefficient, therefore they should be at a +1 level to produce maximum response.

