---
title: "HW Week 11"
author: "Md Ariful Haque Miah, Ayodegi"
date: "11/13/2022"
output:
  html_document:
    df_print: paged
---
Problem 6.8
```{r}
time<-c(rep(-1, 12),rep(1, 12) )
cul_med<-c(rep(-1, 2), rep(1, 2))
obs <- c(21,22,25,26,23,28,24,25,20,26,29,27,37,39,31,34,38,38,29,33,35,36,30,35)
time <- as.factor(time)
cul_med <- as.factor(cul_med)
dat <- data.frame(time,cul_med,obs)
dat
model <- aov(obs~time*cul_med,data=dat)
summary(model) 
plot(model)
```
Hypothesis Test:

$H_0: \alpha_i = 0$ 

$H_a: \alpha_i \neq 0$

$H_0: \beta_j\ = 0$ 

$H_a: \beta_j \neq 0$

$H_0: \alpha\beta_{ij}=0$

$H_a: \alpha\beta_{ij}\neq0$

Only the factors Time and the interaction effects (Time:Culture Medium) are significant as the p value is less than $\alpha=0.001$
level of significance so for these factors we reject $H_0$ and the factor Culture medium is not significant as the p value is larger
than $\alpha=0.05$ level of significance so for this factors we fail to reject $H_0$.
Also Residual vs Fitted plot shows that the variance is constant and the errors are normally distributed and thus the model is adequate.

Problem 6.12
(a) Estimate the factor effects
```{r}
A <- c(rep(-1,4),rep(1,4))
B <- c(rep(-1,8),rep(1,8))
obs <- c(14.037,16.165,13.972,13.907,13.880,13.860,14.032,13.914,14.821,14.757,14.843,14.878,14.888,14.921,14.415,14.932)
A <- as.factor(A)
B <- as.factor(B)
dat <- data.frame(A,B,obs)
dat
```
Finding Corner Points:
```{r}
one <- sum(dat$obs[1:4])
one
a <- sum(dat$obs[5:8])
a
b <- sum(dat$obs[9:12])
b
ab <- sum(dat$obs[13:16])
ab
n <- c(4)
fact_A <- 2*(a+ab-b-one)/(4*n)
fact_A
fact_B <- 2*(b+ab-a-one)/(4*n)
fact_B
fact_AB <- 2*(ab+one-a-b)/(4*n)
fact_AB
```
Therefore, The factor Effect A = -0.31725, B = 0.586, Interaction effect AB = 0.2815

(b) Analysis of variance
```{r}
A <- c(rep(-1,4),rep(1,4))
B <- c(rep(-1,8),rep(1,8))
obs <- c(14.037,16.165,13.972,13.907,13.880,13.860,14.032,13.914,14.821,14.757,14.843,14.878,14.888,14.921,14.415,14.932)
A <- as.factor(A)
B <- as.factor(B)
dat <- data.frame(A,B,obs)
dat
model <- aov(obs~A*B,data=dat)
summary(model)
```
Hypothesis Test: $H_0: \alpha_i = 0$ 

$H_a: \alpha_i \neq 0$

$H_0: \beta_j\ = 0$ 

$H_a: \beta_j \neq 0$

$H_0: \alpha\beta_{ij}=0$

$H_a: \alpha\beta_{ij}\neq0$


p values of the main effect A and the interaction effect (A:B) is greater than $\alpha = 0.05$ level of significance.
Hence the main effect A and the interaction effect (A:B) are not significant as we fail to reject $H_0$.
On the other hand p value of the main effect B is less than $\alpha = 0.1$ level of significance. Hence the main effect B
is significant as we reject $H_0$.

(c) Regression Equation:
```{r}
A <- c(rep(-1,4),rep(1,4))
B <- c(rep(-1,8),rep(1,8))
obs <- c(14.037,16.165,13.972,13.907,13.880,13.860,14.032,13.914,14.821,14.757,14.843,14.878,14.888,14.921,14.415,14.932)
A <- as.factor(A)
B <- as.factor(B)
dat <- data.frame(A,B,obs)
dat
model <- lm(obs~A+B+A*B,data=dat)
summary(model)
```
y = $\beta_0 + \beta_2 *x_2 + \epsilon$

Hence, y = $14.5203 + 0.3045 * x_2 + \epsilon$

where $\beta_0$ = the intercept = the grand mean of all 16 observations and the regression coefficient $\beta_2$ is one-half the
corresponding factor effect B estimate.

(d) The Residuals Analysis
```{r}
A <- c(rep(-1,4),rep(1,4))
B <- c(rep(-1,8),rep(1,8))
obs <- c(14.037,16.165,13.972,13.907,13.880,13.860,14.032,13.914,14.821,14.757,14.843,14.878,14.888,14.921,14.415,14.932)
A <- as.factor(A)
B <- as.factor(B)
dat <- data.frame(A,B,obs)
dat
model <- aov(obs~A*B,data=dat)
summary(model)
plot(model)
```
There is a significant concern in the Residuals vs Fitted Plot, which shows that the variance is not constant. Hence the model is not adequate. Though the normality assumptions are fairly satisfied in the Normal Q-Q plot.

(e) The potential Outlier: Only the observation no. 2 might be the potential outlier according to the $\sqrt$(standardized residual) plot in part (d). Since There is 15 more observations left so only one outlier won't cause any effect to grow an epitaxial layer on polished silicon wafers.

Problem 6.21
(a)
```{r}
A <- c(rep(-1,7),rep(1,7))
B <- c(rep(-1,14),rep(1,14))
C <- c(rep(-1,28),rep(1,28))
D <- c(rep(-1,56),rep(1,56))
obs <- c(10,18,14,12.5,19,16,18.5,0,16.5,4.5,17.5,20.5,17.5,33,4,6,1,14.5,12,14,5,0,10,34,11,25.5,21.5,0,0,0,18.5,19.5,16,15,11,
         5,20.5,18,20,29.5,19,10,6.5,18.5,7.5,6,0,10,0,16.5,4.5,0,23.5,8,8,8,4.5,18,14.5,10,0,17.5,6,19.5,18,16,5.5,10,7,
         36,15,16,8.5,0,0.5,9,3,41.5,39,6.5,3.5,7,8.5,36,8,4.5,6.5,10,13,41,14,21.5,10.5,6.5,0,15.5,24,16,0,0,0,4.5,1,4,6.5,18,
         5,7,10,32.5,18.5,8)
A <- as.factor(A)
B <- as.factor(B)
C <- as.factor(C)
D <- as.factor(D)
dat <- data.frame(A,B,C,D,obs)
dat
model <- aov(obs~A*B*C*D,data=dat)
summary(model)
```
Hypothesis Test:

$H_0: \alpha_i = 0$

$H_a: \alpha_i \neq 0$

$H_0: \beta_j\ = 0$ 

$H_a: \beta_j \neq 0$

$H_0: \gamma_k\ = 0$ 

$H_a: \gamma_k \neq 0$

$H_0: \delta_l\ = 0$ 

$H_a: \delta_l \neq 0$

$H_0: \alpha\beta_{ij}=0$

$H_a: \alpha\beta_{ij}\neq0$

$H_0: \alpha\gamma_{ik}=0$

$H_a: \alpha\gamma_{ik}\neq0$

$H_0: \alpha\delta_{il}=0$

$H_a: \alpha\delta_{il}\neq0$

.....

.....

.....

$H_0: \alpha\beta\gamma\delta_{ijkl}=0$

$H_a: \alpha\beta\gamma\delta_{ijkl}\neq0$

From the anova model summary and analyzing the p value we see that None of the interaction terms are significant but the main effect factors "length of putt" and "the type of putter" are significant and these factors significantly affect putting performance.

Problem 6.21
(b)
```{r}
plot(model)
```
Residual vs fitted plot shows that the variance is not constant and hence the model is not adequate though the errors are assumed to be normally distributed from the normal Q-Q plot.

