---
title: "HW13"
author: "Ayodeji, Md Ariful Haque"
date: '2022-11-25'
output: html_document
---

## Question 8.2

We are experimenting on $2^{4-1}$ where k=4 and p=1

I=ABCD and I=-ABCD

```{r,warning=FALSE,message=FALSE}
library(FrF2)
res3<-FrF2(nfactors=4,resolution=4,randomize=FALSE)
res3
```

```{r}
response<- c(7.037,16.867,13.876,17.273,11.846,4.368,9.36,15.653)
response1 <-  add.response(res3,response)
aliasprint(response1)
```

We can conclude that since the two factor interaction is aliased with each other, they are in the $4^{th}$ resolution.

Running the summary of our design we have

```{r}
summary(response1)
```

```{r}
DanielPlot(response1)
MEPlot(response1,show.alias=TRUE)
```

We can see from the Daniel plot that none of the factors are significant.

Because of this we can say that none of the factors affect the Crack length.

## Question 8.24

We are experimenting on $2^{5-1}$ where k=5 and p=1

I=ABCDE and I=-ABCDE

```{r}
res1<- FrF2(nfactors = 5, resolution = 5 ,randomize = FALSE)
aliasprint(res1)
```

The summary of our design is

```{r}
summary(res1)
```

Confounding the AB factor combination with blocks

```{r}
AB<- c("+","-","-","+","+","-","-","+","+","-","-","+","+","-","-","+")
block<- c(1,2,2,1,1,2,2,1,1,2,2,1,1,2,2,1)
dat<- data.frame(res1,AB,block)
dat
```

We can see that the AB and CDE factor interactions are confounded with blocks

Also, we see that No Two factor interaction is confounded in block 1 and block 2

Overall we noticed that,

Main effects confounded in block 1- E, C,D

Main effects confounded in block 2- A,B

## Question 8.25

We are experimenting on $2^{7-2}$ where k=7 and p=2

I=ABCDEF and I=-ABCDEF ;

I=ABDEG and I=-ABDEG

With a generalized interaction of I=CEFG, I=-CEFG

```{r}
design1<- FrF2(nruns = 32,nfactors=7,blocks = 4,randomize=TRUE)
design1
```

```{r}
summary(design1)


```

We observe that only ACE, BFG, and ABCEFG are confounded with blocks

## Question 8.28

```{r}
Lamtemp <- c(rep(c("-1","1"),8)) 
Lamtime <- c(rep(c("-1","-1","1","1"),4))
LamPres<- c(rep(c("-1","-1","-1","-1","1","1","1","1"),2))
Firtemp <- c(rep(c("-1","-1","-1","-1","-1","-1","-1","-1","1","1","1","1","1","1","1","1"),1))
Fircytime <- c(rep(c("-1","1","1","-1","1","-1","-1","1"),2))
Firdewpoint <- c(rep(c("-1","1","-1","1","1","-1","1","-1"),2))
response<- c(0.0167,0.0062,0.0041,0.0073,0.0047,0.0219,0.0121,0.0255,0.0032,0.0078,0.0043,0.0186,0.011,0.0065,0.0155,0.0093,0.0128,0.0066,0.0043,0.0081,0.0047,0.0258,0.009,0.025,0.0023,0.0158,0.0027,0.0137,0.0086,0.0109,0.0158,0.0124,0.0149,0.0044,0.0042,0.0039,0.004,0.0147,0.0092,0.0226,0.0077,0.006,0.0028,0.0158,0.0101,0.0126,0.0145,0.011,0.0185,0.002,0.005,0.003,0.0089,0.0296,0.0086,0.0169,0.0069,0.0045,0.0028,0.0159,0.0158,0.0071,0.0145,0.0133)
total<- c(629,192,176,223,223,920,389,900,201,341,126,640,455,371,603,460)
mean<- c(157.25,48,44,55.75,55.75,230,97.25,225,50.25,85.25,31.5,160,113.75,92.75,150.75,115)
std <- c(24.418,20.976,4.083,25.025,22.41,63.639,16.029,39.42,26.725,50.341,7.681,20.083,31.12,29.51,6.75,17.45
)
dat<-cbind(Lamtemp,Lamtime,LamPres,Firtemp,Fircytime,Firdewpoint,response,total,mean,std)
dat <-as.data.frame(dat)
dat
```

Question a.

The above design uses a $2^{6-2}$ with Resolution IIII and with 16 runs.

## b.)

the following which are alias relationship are displayed below

```{r}
res4<- FrF2(nfactors = 6,resolution = 4 , randomize = TRUE)
aliasprint(res4)
```

Using table 8.14 from the Montgomery textbook,

The design generators is

E=(+-ABC) and F=(+-BCD) that is I=ABCE , I=BCDF

The generalized interaction is I=ADEF

### C.

```{r}
Lamtemp <- c(rep(c(-1,1),8)) 
Lamtime <- c(rep(c(-1,-1,1,1),4))
LamPres <- c(rep(c(-1,-1,-1,-1,1,1,1,1),2))
Firtemp <- c(rep(c(-1,-1,-1,-1,-1,-1,-1,-1,1,1,1,1,1,1,1,1),1))
Fircytime <- c(rep(c(-1,1,1,-1,1,-1,-1,1),2))
Firdewpoint <- c(rep(c(-1,1,-1,1,1,-1,1,-1),2))
response<- c(0.0167,0.0062,0.0041,0.0073,0.0047,0.0219,0.0121,0.0255,0.0032,0.0078,0.0043,0.0186,0.011,0.0065,0.0155,0.0093,0.0128,0.0066,0.0043,0.0081,0.0047,0.0258,0.009,0.025,0.0023,0.0158,0.0027,0.0137,0.0086,0.0109,0.0158,0.0124,0.0149,0.0044,0.0042,0.0039,0.004,0.0147,0.0092,0.0226,0.0077,0.006,0.0028,0.0158,0.0101,0.0126,0.0145,0.011,0.0185,0.002,0.005,0.003,0.0089,0.0296,0.0086,0.0169,0.0069,0.0045,0.0028,0.0159,0.0158,0.0071,0.0145,0.0133)
total<-c(629,192,176,223,223,920,389,900,201,341,126,640,455,371,603,460)
mean<- c(157.25,48,44,55.75,55.75,230,97.25,225,50.25,85.25,31.5,160,113.75,92.75,150.75,115)
std <- c(24.418,20.976,4.083,25.025,22.41,63.639,16.029,39.42,26.725,50.341,7.681,20.083,31.12,29.51,6.75,17.45
)
dat <- cbind(Lamtemp,Lamtime,LamPres,Firtemp,Fircytime,Firdewpoint,response,total,mean,std)
dat <- as.data.frame(dat)
dat
```

```{r,message=TRUE,warning=FALSE}
library(GAD)
dat$Lamtemp <- as.fixed(dat$Lamtemp)
dat$Lamtime <- as.fixed(dat$Lamtime)
dat$LamPres <- as.fixed(dat$LamPres)
dat$Firtemp <- as.fixed(dat$Firtemp)
dat$Firdewpoint <- as.fixed(dat$Firdewpoint)
dat$Fircytime <- as.fixed(dat$Fircytime)

model<- aov(response~Lamtemp*Lamtime*LamPres*Firtemp*Fircytime*Firdewpoint,data = dat)
summary(model)

```

The variables that would affect average camber is A,C,E,F also some interaction effects that can be seen above.

## D.

```{r}
camber<-  c(24.418,20.976,4.083,25.025,22.41,63.639,16.029,39.42,26.725,50.341,7.681,20.083,31.12,29.51,6.75,17.45)
var<-c(camber^2)
A<- c(rep(c("-1","1"),8)) 
B<- c(rep(c("-1","-1","1","1"),4))
C<- c(rep(c("-1","-1","-1","-1","1","1","1","1"),2))
D<- c(rep(c("-1","-1","-1","-1","-1","-1","-1","-1","1","1","1","1","1","1","1","1"),1))
E<- c(rep(c("-1","1","1","-1","1","-1","-1","1"),2))
F<- c(rep(c("-1","1","-1","1","1","-1","1","-1"),2))
dat2<- data.frame(A,B,C,D,E,F,var)
model2<-lm(camber~A*B*C*D*E*F,data=dat2)
DanielPlot(model2)
```

We conclude that based on the Daniel's plot that

Laminating temperature and Laminating time are factors that are significantly affecting the standard deviation.

Running the model of the significant factors above we have

```{r}
model5<-aov(camber~A+B,data = dat2)
summary(model5)
```

From the ANOVA analysis, we can see that since the pvalue of Laminating temperature(0.01202) and Laminating time(0.00948) are less than the reference level of significance(0.05).

We ultimately conclude that Laminating temperature and Laminating time are affecting standard deviation of camber.

## E.

```{r}
Lamtemp <- c(rep(c("-1","1"),8)) 
Lamtime <- c(rep(c("-1","-1","1","1"),4))
LamPres <- c(rep(c("-1","-1","-1","-1","1","1","1","1"),2))
Firtemp <- c(rep(c("-1","-1","-1","-1","-1","-1","-1","-1","1","1","1","1","1","1","1","1"),1))
Fircytime <- c(rep(c("-1","1","1","-1","1","-1","-1","1"),2))
Firdewpoint <- c(rep(c("-1","1","-1","1","1","-1","1","-1"),2))
response<- c(0.0167,0.0062,0.0041,0.0073,0.0047,0.0219,0.0121,0.0255,0.0032,0.0078,0.0043,0.0186,0.011,0.0065,0.0155,0.0093,0.0128,0.0066,0.0043,0.0081,0.0047,0.0258,0.009,0.025,0.0023,0.0158,0.0027,0.0137,0.0086,0.0109,0.0158,0.0124,0.0149,0.0044,0.0042,0.0039,0.004,0.0147,0.0092,0.0226,0.0077,0.006,0.0028,0.0158,0.0101,0.0126,0.0145,0.011,0.0185,0.002,0.005,0.003,0.0089,0.0296,0.0086,0.0169,0.0069,0.0045,0.0028,0.0159,0.0158,0.0071,0.0145,0.0133)
total <- c(629,192,176,223,223,920,389,900,201,341,126,640,455,371,603,460)
mean <- c(157.25,48,44,55.75,55.75,230,97.25,225,50.25,85.25,31.5,160,113.75,92.75,150.75,115)
std <- c(24.418,20.976,4.083,25.025,22.41,63.639,16.029,39.42,26.725,50.341,7.681,20.083,31.12,29.51,6.75,17.45
)
dat3<- cbind(Lamtemp,Lamtime,LamPres,Firtemp,Fircytime,Firdewpoint,response,total,mean,std)
dat3<- as.data.frame(dat3)
model5<-lm(dat$response~dat$Lamtemp*dat$Lamtime*dat$LamPres*dat$Firtemp*dat$Fircytime*dat$Firdewpoint)
coef(model5)
```

```{r}
summary(model5)
```

The expression for reducing camber(y) is

y= 0.015735 + 0.0010094A +0.0017844C -0.0041875E - 0.0077469F

In conclusion we have to let A, C be at low levels [-1,-1] and E,F high [+1,+1].

## Question 8.40

Part A,

the experiment investigated A,B,C,D which are four factors.

Part B,

the resolution of the design used is **FOUR**

Part C,

```{r}
O = 8
AD =10
BD =12
AB =7
CD =13
AC = 6
BC =5
ABCD =11
EffectofA<-(2*(AD+AB+AC+ABCD-O-BD-CD-BC))/(8)
EffectofA
```

```{r}
EffectofB<-(2*(BD+AB+BC+ABCD-O-AD-CD-AC))/(8)
EffectofB
```

```{r}
EffectofC <- (2*(CD+AC+BC+ABCD-O-AD-BD-AB))/(8)
EffectofC
```

```{r}
EffectofD <- (2*(AD+BD+CD+ABCD-O-AB-AC-BC))/(8)
EffectofD
```

We can see that

Effect of A is -1

Effect of B is -0.5

Effect of C is -0.5

Effect of D is 5

## QUESTION 8.48.

```{r}
design4<- FrF2(nfactors = 5,nruns = 8,generators = c("-ABC","BC"), randomize = FALSE)
design4
```

```{r}
summary(design4)
```

```{r}
design5<-fold.design(design4)
aliasprint(design5)
```

a\. the design generator for column D is -ABC

b\. the design generator for column E is BC

c\. From the Montgometry textbook the resolution of the design is **FOUR.**

## Question 8.60

```{r,message=FALSE,warning=FALSE}
library(FrF2)
design<- FrF2(nfactors=7,resolution=3,randomize=FALSE)
design
```

```{r}
design2<-fold.design(design,column=1)
design2
```

```{r}
design.info(design2)$aliased$main

```

```{r}
aliasprint(design2)
```
