---
title: "Homework10"
author: "Ayodeji, Md Ariful Haque"
date: '2022-11-05'
output: html_document
---

## Question 5.4

```{r,message=FALSE,warning=FALSE,error=FALSE}
Feedrate<- c(rep(1,12), rep(2,12), rep(3,12))
cut<- rep(seq(1,4),9)
obs <- c(74, 79, 82, 99, 64, 68, 88, 104, 60, 73, 92, 96, 92, 98, 99, 
              104, 86, 104, 108, 110, 88, 88, 95, 99, 99, 104, 108, 114, 
              98, 99, 110, 111, 102, 95, 99, 107)
dat<- data.frame(Feedrate,cut,obs)
```

Question 5.4 a

Testing at level of significance=0.05

Using the model equation of

$Yijk$=$\mu$+$\alpha i$+$\beta j$+$\alpha\beta ij$+$\varepsilon ijk$

where $\varepsilon ijk$ $\sim \left( 0,\sigma^{2} \right)$

Stating our null and alternative hypothesis we have

Null hypothesis for interaction effects are

$\alpha\beta ij=0$ for all ij

Alternative hypothesis for interaction effects are

$\alpha\beta ij\neq 0$ for some ij

Main effect hypothesis

For Null hypothesis

$\alpha i=0$ for all i

for alternative hypothesis

$\alpha i\neq 0$ for some i

Null hypothesis

$\beta j=0$ for all j

$\beta j\neq 0$ for some j

```{r,message=FALSE,error=FALSE,warning=FALSE}
library(GAD)
Feedrate<- as.fixed(Feedrate)
cut <- as.fixed(cut)
Model <- aov(obs~cut+Feedrate+cut*Feedrate)
GAD::gad(Model)
```

From the values obtained from our model we can see that

the p-value for the interaction model (**0.01797**) is less than our reference significant level of alpha(**0.05**) , hence we are stating that we are rejecting the null hypothesis and saying that there is interaction between the feedrate and depth of cut in our analysis

The p-value for our feed-rate was highly significant with p=1.086e-09, which was less than 0.05

The p-value for our depth of cut was also highly significant with p=1.652e-07, which was less than 0.05

## 

```{r}
interaction.plot(cut,Feedrate,obs)
```

The above interaction plot concludes that we have an interaction effects between the two factors considered (depth of cut and feed-rate)

## 

## Question 5.4 b

```{r}
plot(Model)
```

From the residual plot, we can see that the Normal probability plot of the residuals falls fairly on a straight line and hence we can validate normality

Since the residual vs fitted plot falls fairly on a straight line ,we can roughly assume constant variance in our model

## Question 5.4 c

```{r}
mean(obs[1:12])
mean(dat$obs[1:12])
```

```{r}
var(obs[1:12])
var(dat$obs[1:12])
```

```{r}
mean(dat$obs[13:24])
```

```{r}
var(dat$obs[13:24])
```

```{r}
mean(dat$obs[25:36])
```

```{r}
var(dat$obs[25:36])
```

point estimate for feed-rate 0.2

mean= 81.58 and variance=205.53

point estimate for feed-rate 0.25

mean= 97.58 and variance=64.08

point estimate for feed-rate 0.3

mean=103.83 and variance=36.87

Part D

Find the P-values for model in part (a).

P-value of Depth of Cut & Feedrate Interaction: 0.01797(significant)

P-value of Depth of Cut : 1.652e-07 (significant)

P-value of Feedrate : 1.086e-09(significant)

## QUESTION 5.34

Using the model equation of

$Yijkl$=$\mu$+$\alpha i$+$\beta j$+$\gamma k$+$\alpha\beta ij$+$\varepsilon ijk$

$\gamma k$=block effect

where $\varepsilon ijk$ $\sim \left( 0,\sigma^{2} \right)$

Stating our null and alternative hypothesis we have

Null hypothesis for interaction effects are

$\alpha\beta ij=0$ for all ij

Alternative hypothesis for interaction effects are

$\alpha\beta ij\neq 0$ for some ij

Main effect hypothesis

For Null hypothesis

$\alpha i=0$ for all i

for alternative hypothesis

$\alpha i\neq 0$ for some i

Null hypothesis

$\beta j=0$ for all j

$\beta j\neq 0$ for some j

```{r}
Feedrate<- c(rep(1,12), rep(2,12), rep(3,12))
cut<- rep(seq(1,4),9)
block<- c(rep(1,4),rep(2,4),rep(3,4),rep(1,4), rep(2,4),rep(3,4),rep(1,4),rep(2,4),rep(3,4))
obs<- c(74, 79, 82, 99, 64, 68, 88, 104, 60, 73, 92,96, 92, 98, 99,104, 86, 104, 108, 110, 88, 88, 95,99, 99, 104, 108, 114, 98, 99, 110, 111, 102, 95, 99, 107)
dat<- data.frame(cut,Feedrate,block,obs)
```

```{r}
library(GAD)
Feedrate<- as.fixed(Feedrate)
cut <- as.fixed(cut)
block <- as.fixed(block)
Model2<- aov(obs~cut+Feedrate+block+cut*Feedrate)
summary(Model2)
```

Checking out the p-value(0.00726\<0.05) of the interaction effects, we can see there exists an interaction effect between the two factors.

We can also see that the p-values of the main effects are also significant

p-value of depth of cut - 4.89e-08 (highly significant at alpha=0.05)

p-value of feed-rate - 3.64e-10 ((highly significant at alpha=0.05)

## The variance components of blocks is

Msb=90.3

Mse=23.1

i=3 , j=4

i\*j=12

```{r}
var<- (90.3-23.1)/12
var
```

The variance of the block= 5.6

from our model the p-value of blocks is 0.03532 which is less than our reference significant level of alpha(0.05), we can confidently state that blocking had a significant effect in this case and therefore we conclude that it was important to block.

## Interaction plot

```{r}
interaction.plot(Feedrate,cut,obs)
```

The above interaction plot concludes that we have an interaction effects between the two factors considered (depth of cut and feed-rate).

## Question 13.5

Using the model equation of

$Yijk$=$\mu$+$\alpha i$+$\beta j$+$\alpha\beta ij$+$\varepsilon ijk$

where $\varepsilon ijk$ $\sim \left( 0,\sigma^{2} \right)$

Stating our null and alternative hypothesis we have

Null hypothesis for interaction effects are

$\sigma^{2}_{\alpha\beta}=0$

Alternative hypothesis for interaction effects are

$\sigma^{2}_{\alpha\beta}\neq 0$

Main effects (hypothesis)

Null hypothesis

$\sigma^{2}_{\alpha}=0$

Alternative hypothesis

$\sigma^{2}_{\alpha}\neq 0$

Null hypothesis

$\sigma^{2}_{\beta}=0$

Alternative hypothesis

$\sigma^{2}_{\beta}\neq 0$

```{r}
p<- c(rep(1,9), rep(2,9))
t<- rep(seq(1,3),6)
obs<- c(570, 1063, 565, 565, 1080, 510, 583, 1043, 590, 528, 988, 526, 547, 1026,
              538, 521, 1004, 532)
p<- as.random(p)
t<- as.fixed(t)
dat<- data.frame(p,t,obs)
```

```{r}
library(GAD)
model3<- aov(obs~p+t+p*t,data=dat)
GAD::gad(model3)
```

Checking out the p-value(0.4271101 \>0.05) of the interaction effects(position and temperature), we can see there exists no interaction effect between the two factors.(we are failing to reject the null hypothesis)

But

We can also that the p-values of the main effects are all significant

p-value of position - 0.0017624 (highly significant at alpha=0.05)

p-value of temperature - 0.0008647 (highly significant at alpha=0.05)

## Question 13.6

Using the model equation of

$Yijk$=$\mu$+$\alpha i$+$\beta j$+$\alpha\beta ij$+$\varepsilon ijk$

where $\varepsilon ijk$ $\sim \left( 0,\sigma^{2} \right)$

Stating our null and alternative hypothesis we have

Null hypothesis for interaction effects are

$\sigma^{2}_{\alpha\beta}=0$

Alternative hypothesis for interaction effects are

$\sigma^{2}_{\alpha\beta}\neq 0$

Main effects (hypothesis)

Null hypothesis

$\sigma^{2}_{\alpha}=0$

Alternative hypothesis

$\sigma^{2}_{\alpha}\neq 0$

Null hypothesis

$\sigma^{2}_{\beta}=0$

Alternative hypothesis

$\sigma^{2}_{\beta}\neq 0$



```{r}
parts<- c(rep(1,6), rep(2,6), rep(3,6), rep(4,6), rep(5,6), rep(6,6), rep(7,6), rep(8,6), rep(9,6), rep(10,6))
operators<- c(rep(1,3), rep(2,3), rep(1,3), rep(2,3),rep(1,3), rep(2,3), rep(1,3),rep(2,3), rep(1,3), rep(2,3), rep(1,3),rep(2,3), rep(1,3), rep(2,3),rep(1,3), rep(2,3), rep(1,3), rep(2,3), rep(1,3), rep(2,3))
obs<- c(50, 49, 50, 50, 48, 51, 52, 52, 51, 51, 51, 51, 53, 50, 50, 54, 52, 51,49, 51, 50, 48, 50, 51, 48, 49, 48, 48, 49, 48, 52, 50, 50, 52, 50, 50,51, 51, 51, 51, 50, 50, 52, 50, 49, 53, 48, 50, 50, 51, 50, 51, 48, 49,47, 46, 49, 46, 47, 48)
parts<-as.random(parts)
operators<- as.fixed(operators)
dat3<- data.frame(parts,operators,obs)
dat3
```

```{r}
library(GAD)
model4<- aov(obs~parts*operators,data=dat3)
gad(model4)
interaction.plot(operators,parts,obs)

```

Checking out the p-value (0.9270\>0.05) of the interaction effects(operator and parts), we can see there exists no interaction effect between the two factors.(we are failing to reject the null hypothesis)

But

p-value of parts - 3.216e-06 (highly significant at alpha=0.05)

p-value of operator - 0.4269 (insignificant at alpha=0.05)
